---
title: "DAG and identification"
author: "Yangyong Ye"
format: revealjs
editor: visual
execute: 
  echo: false
---

## DAG

-   Displays **assumptions** about the relationship between variables (**nodes**).

-   The assumptions take the form of lines (**edges**) going from one node to another.

-   Directed edges means that they have a single arrowhead indicating their effect.

-   No assumptions on functional form or variable distribution.

```{r}
library(ggdag)
library(ggplot2)
theme_set(theme_dag())
dagify(y~x) %>% 
  ggdag()
```

## bi-direction

```{r}
dagify(y ~~ x) %>% 
  ggdag()
```

## bi-direction = confounder

```{r}
dagify(y ~~ x) %>% 
  ggdag_canonical() 
```

## acyclic

```{r}
dagify(y ~ x,
       x ~ a,
       a ~ y) %>% 
  ggdag() 

```

## mediation

```{r}

ggdag_mediation_triangle()

```

## confounder

```{r}
ggdag_confounder_triangle()

```

## collider

```{r}

ggdag_collider_triangle()
```

## collider

```{r}
fever_dag <- collider_triangle(x = "Influenza", 
                  y = "Chicken Pox", 
                  m = "Fever") 

ggdag(fever_dag, text = FALSE, use_labels = "label")
```

## collider, control or not control

```{r}
ggdag_dseparated(fever_dag, text = FALSE, use_labels = "label")
```

## collider, control or not control

```{r}
ggdag_dseparated(fever_dag, controlling_for = "m", 
                 text = FALSE, use_labels = "label")
```

## descendant of colliders

```{r}
dagify(fever ~ flu + pox, 
        acetaminophen ~ fever,
        labels = c("flu" = "Influenza",
                   "pox" = "Chicken Pox",
                   "fever" = "Fever",
                   "acetaminophen" = "Acetaminophen")) %>% 
ggdag_dseparated(from = "flu", to = "pox", controlling_for = "acetaminophen",
                 text = FALSE, use_labels = "label")
```

## mediation

```{r}
smoking_ca_dag <- dagify(cardiacarrest ~ cholesterol,
       cholesterol ~ smoking + weight,
       smoking ~ unhealthy,
       weight ~ unhealthy,
       labels = c("cardiacarrest" = "Cardiac\n Arrest", 
                  "smoking" = "Smoking",
                  "cholesterol" = "Cholesterol",
                  "unhealthy" = "Unhealthy\n Lifestyle",
                  "weight" = "Weight"),
       latent = "unhealthy",
       exposure = "smoking",
       outcome = "cardiacarrest")
ggdag_dseparated(smoking_ca_dag, controlling_for = c("weight", "cholesterol"), 
                 text = FALSE, use_labels = "label", collider_lines = FALSE)
```

## complicated

```{r}

dag1 <- dagify(Y ~ D + I,
               D ~ B + I + PE,
               I ~ PE,
               PE ~ B,
               exposure = "D",
               outcome = "Y",
               latent = "B")

ggdag(dag1)


```

## parents

```{r}
ggdag_parents(dag1, "D")
```

## 

```{r}
ggdag_children(dag1, "D")
```

## ancestors

```{r}
ggdag_ancestors(dag1, "Y")
```

## descendants

```{r}
ggdag_descendants(dag1,"B")
```

## path

```{r}
ggdag_paths(dag1,shadow = T)

```

## adjustment set

```{r}
ggdag_adjustment_set(dag1, exposure = "D",
                     outcome = "Y")
```

## collider bias

```{r}
dag2 <- dagify(Y ~ D + U1 + U2,
               D ~ U1 + U2,
               I ~ U1 + U2,
               exposure = "D",
               outcome = "Y",
               latent = c("U1","U2"))
ggdag(dag2)
```

## 

```{r}
ggdag_paths(dag2,shadow = T)
```

## adjustment

```{r}
ggdag_adjustment_set(dag2)
```
